---
title: "Hematocrit"
author: "Elliott Schmidt"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    keep_md: yes
    code_folding: show
    collapse: no
    df_print: paged
    fig_caption: yes
    fig_height: 4
    fig_width: 6
    highlight: monochrome
    theme: flatly
    latex_engine: xelatex
    toc: yes
    toc_float: yes
    css: styles.css
  pdf_document:
    df_print: default
    fig_caption: yes
    fig_height: 4
    fig_width: 4
    highlight: tango
    latex_engine: xelatex
    number_sections: yes
    toc_depth: 2
documentclass: article
fontsize: 12pt
mainfont: Arial
mathfont: LiberationMono
classoption: a4paper
--- 

# Scenario 

For initial details on the experiment performed please read the **ReadMe** file. In breif, _Acanthochromis polyacanthus_ from two different regions on the Great Barrier Reef (GBR) were tested for metabolic performance at four different temperatures, 27$^\circ$C, 28.5$^\circ$C, 30$^\circ$C, and 31.5$^\circ$C. Fish used in this study were collected from two different regions, low- (i.e. Cairns) and high-latitude (i.e., Mackay), within each region fish were collected from a total of three different populations. 

Blood samples for hematocrit sampling were collected 2-weeks after fish underwent  respiormetry testing at the final experimental temperature (31.5$^\circ$C). Hematocrit ratios were measured by comparing the amount of packed red blood cells to blood plasma, after blood samples collected via capillary tubes. 


# Read in the data

Before beginning always make sure that you are working in the correct directory 

```{r set-dir-hide, include=FALSE}
working.dir = "C:/Users/jc527762/OneDrive - James Cook University/PhD dissertation/Data/Chapter1_LocalAdaptation/"
```

```{r set_dir}
knitr::opts_knit$set(root.dir=working.dir)
```

Lets start by loading the packages that are needed 

## Load packages 

```{r load-packages, warning=FALSE, message=FALSE}
library(tidyverse) # data manipulation
library(plyr) # data manipulation
library(dplyr) # data manipulation
library(lubridate) # data manipulation - specifically time data
library(ggplot2) # plotting figures
library(glmmTMB) # running models
library(performance) # model validation
library(car) # dependent
library(DHARMa) # model validation
library(MuMIn) # model validation
library(kableExtra) # creating tables
library(broom) # dependent
library(emmeans) # post-hoc analysis
library(ggeffects) # plotting models/model validation
library(vtable) # creating tables
library(modelr) # model validation
library(kableExtra) # formatting output tables
library(sjPlot) # plotting models
```



Now we can import that data. Replace import data with the PATH to your data file. I have secretly labelled my PATH import.data (i.e. import.data = "PATH TO MY FILE")

## Load data 
```{r import-data-hide, include=FALSE}
import.data = read.delim("C:/Users/jc527762/OneDrive - James Cook University/PhD dissertation/Data/Local_adaptation/Chapter1_LocalAdaptation/import_files/HematocritHemoglobin.txt")
```

```{r load-data}
hema <- import.data
``` 

# Data manipulation 

Before the data can be analysed it is important to clean up the data file. Below a number of adjustments are made, primarily making sure that columns are being treated appropriately as either factors, numeric, or as time, as well as the renaming of some columns.  

```{r data-manipulation-1, warning=FALSE, message=FALSE, class.source="style"}
hema <-  hema %>% 
  mutate(PERC_RBC = as.numeric(PERC_RBC), 
         MASS = as.numeric(MASS),
         MASS_CENTERED = scale(MASS, center = TRUE, scale = FALSE)) %>% 
  drop_na(PERC_RBC)
```

Great! That is everything for data manipulation 

# Exploratory data analysis {.tabset}

## HEMATOCRIT V LATITUDE

```{r eda-1, warning=FALSE, message=FALSE}
ggplot(hema, aes(y=PERC_RBC, x=REGION)) + 
  geom_boxplot() + 
  theme_classic() 
```

## HEMATOCRIT V LATITUDE (distr)
```{r eda-2, warning=FALSE, message=FALSE}
hema %>% ggplot(aes(x=PERC_RBC)) + 
  geom_density() +
  facet_wrap(~REGION)
```

## {-}

# Fit the model 

The model was fit using the **glm** and later **glmmTMB** package in R. A number of different models were tested to determine which hypothesis and associated variables best predicted resting oxygen consumption. Model fit was examined using AICc, BIC, and r-squared values. Additional model were examined via the validation diagnostics provided by the **performance** and **dHARMA** packages in R. 

## Fixed factors (linear regression models)

### model 1
```{r model-fit-1, warning=FALSE}
#--- base model ---#
hema.1 <- glm(PERC_RBC ~ REGION, 
                family = gaussian(),  
                data = hema) 
```
#### summary
```{r model-fit-1.2, warning=FALSE, echo=FALSE}
as.data.frame(summary(hema.1)$coefficients) %>% 
  knitr::kable(format = "html")  %>%
  kable_paper(full_width = TRUE)

```

### model 2
```{r model-fit-2, warning=FALSE}
#--- experimental rmr equipment hypothesis ---#
hema.2 <- glm(PERC_RBC ~ REGION + MASS_CENTERED, 
                 family = gaussian(), 
                 data = hema)  
```

#### summary
```{r model-fit-2.2, warning=FALSE, echo=FALSE}
as.data.frame(summary(hema.2)$coefficients) %>% 
  knitr::kable(format = "html")  %>%
  kable_paper(full_width = TRUE)

```

### model comparison table
```{r model-fit-3, warning=FALSE, echo=FALSE}
aic.c=AICc(hema.1, hema.2, k=2) 
bic=BIC(hema.1, hema.2)

model.fit.table <- as.data.frame(aic.c) %>% 
  tibble::rownames_to_column("model") %>%
  mutate(BIC = bic$BIC, 
         r2 = c(r.squaredGLMM(hema.1)[1], r.squaredGLMM(hema.2)[1]))
model.fit.table %>% knitr::kable(format = "html")  %>%
  kable_paper(full_width = TRUE)
```

There is little difference between the two initial models, however, the model that does not include **MASS_CENTERED** prefers better via the model comparisons scores, therefore, we will move forward with the first and most simple model. 

## Random factors 

Fish were only sampled once, therefore, there is no need to include individual as a random factor. However, we will test how the inclusion of **POPULATION** influences the model.  

### random factor models

```{r random-factors-1}
hema.1 <- glmmTMB(PERC_RBC ~ REGION, 
                   family = gaussian(), 
                   REML = TRUE, 
                   data = hema) 

hema.1b <- glmmTMB(PERC_RBC ~ REGION + (1|POPULATION), 
                    family = gaussian(), 
                    REML = TRUE, 
                    data = hema) 

hema.1c <- glmmTMB(PERC_RBC ~ REGION + (REGION|POPULATION), 
                    family = gaussian(), 
                    REML = TRUE, 
                    data = hema) # convergence problem

```

#### random factor model comparisons 

```{r random-factors-1.2, warning=FALSE, echo=FALSE}
aic.c=AICc(hema.1, hema.1b, hema.1c, k=2)
bic=BIC(hema.1, hema.1b, hema.1c)

model.fit.table <- as.data.frame(aic.c) %>% 
  tibble::rownames_to_column("model") %>%
  mutate(BIC = bic$BIC, 
         r2m = c(r.squaredGLMM(hema.1)[1], r.squaredGLMM(hema.1b)[1], r.squaredGLMM(hema.1c)[1]), 
         r2c = c(r.squaredGLMM(hema.1)[2], r.squaredGLMM(hema.1b)[2], r.squaredGLMM(hema.1c)[2]))
model.fit.table %>% knitr::kable(format = "html")  %>%
  kable_paper(full_width = TRUE)

```

The inclusion of random effects does help explain additional variation and therefore will not be included in the model. Note the final model will be run using **glm** and not **glmmmTMB** because we are not using a mixed model. 



# Model validation {.tabset .tabset-faded}

```{r final-model-1, fig.width=8, fig.height=8, message=FALSE, warning=FALSE, echo=FALSE}
hema.1 <- glm(PERC_RBC ~ REGION, 
                family = gaussian(),  
                data = hema) 
```

## performance {.tabset .tabset-faded}

### hema.1
```{r model-valid-1, fig.width=8, fig.height=8, message=FALSE, warning=FALSE, echo=FALSE}
hema.1  %>% performance::check_model()
```

## DHARMa residuals {.tabset .tabset-faded}

### hema.1 
```{r model-valid-2, fig.width=8, fig.height=8, message=FALSE, warning=FALSE, results='markdown'}
hema.1 %>% simulateResiduals(plot=TRUE)
hema.1 %>% DHARMa::testResiduals(plot=TRUE)
```

## {-}

# {-}

The basic looks good and passes the validation checks. 

# Partial plots {.tabset .tabset-faded}

## ggemmeans 

```{r partial-plots-1, message=FALSE, echo=FALSE, fig.width=8, fig.height=6}
hema.1 %>% ggemmeans(~REGION) %>% plot(add.data=TRUE, jitter=c(0.05,0))
```

## plot_model 

```{r partial-plots-2, echo=FALSE, fig.width=8, fig.height=6}
hema.1 %>% plot_model(type='std', vline.color = "grey12", show.values=TRUE, sort.est = TRUE, transform = NULL, value.offset = 0.3)
```

# {-} 

# Model investigation {.tabset .tabset-faded}

## summary 
```{r model-inv-1, echo=FALSE}
as.data.frame(summary(hema.1)$coefficients) %>%
  knitr::kable(format = "html")  %>%
  kable_paper(full_width = TRUE)
```

## Anova 
```{r model-inv-2, echo=FALSE}
hema.1 %>% Anova() %>% 
  as.data.frame() %>%
  knitr::kable(format = "html")  %>%
  kable_paper(full_width = TRUE)
``` 

## confint 
```{r model-inv-3, echo=FALSE}
hema.1 %>% confint() %>% 
  as.data.frame() %>%
  knitr::kable(format = "html")  %>%
  kable_paper(full_width = TRUE)
``` 

## r-squared
```{r model-inv-4, echo=FALSE}
hema.1 %>% performance::r2(method="marginal") %>% 
  as.data.frame() %>%
  knitr::kable(format = "html")  %>%
  kable_paper(full_width = TRUE)
```


# {-} 

# Pairwise comparisons {.tabset .tabset-faded} 

## emmeans [latitudes]
```{r pairwise-2}
hema.1 %>% emmeans(pairwise ~ REGION, type = "response")
```

# {-}

# Summary figure 

```{r sum-fig, fig.width=8, fig.height=6, echo=FALSE}
hema.newdata <-  hema.1 %>% ggemmeans(~REGION) %>% 
  as.data.frame() %>% 
  dplyr::rename(REGION = x)

obs <- hema %>% 
  mutate(Pred = predict(hema.1, re.form=NA), 
         Resid = residuals(hema.1, type = "response"), 
         Fit = Pred + Resid)

hematocrit.plot <- ggplot(hema.newdata, aes(y=predicted, x=REGION, color=REGION))  + 
  geom_jitter(data=obs, aes(y=Pred, x=REGION, color =REGION), 
              width = 0.05, alpha=0.3)+
  geom_pointrange(aes(ymin=conf.low, 
                      ymax=conf.high), 
                  shape = 19, 
                  size = 1, 
                  position = position_dodge(0.2)) + 
  scale_color_manual(values=c("#B2182B", "#4393C3"), labels = c("Low","High"),
                     name = "Latitude") +
  ylab("HEMATOCRIT RATIO") +
  scale_x_discrete(name = "LATITUDE", 
                   labels = c("Low","High"))+
  theme_classic() + 
  theme(legend.position = 'none')  + 
  annotate("text", x=1.5, y=0.275, fontface="italic", size=5, label="P =0.057"); hematocrit.plot
```

# Conclusion 

* In conclusion there is no significant difference in hematocrit ratios between _A. polyacanthus_ from low- and high-latitude at 31.5$^\circ$C. 
